---
title: "Econ 613 Assignment 2"
author: "Xiaolin Ding"
date: "`r format(Sys.time(), '%m/%d/%y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
library(bayesm)
library(dplyr)
library(tidyverse)
data(margarine)
choiceprice = margarine$choicePrice
demos = margarine$demo
```

# Exercise 1 Data Description

## Average prices of each product
```{r}
print(apply(as.matrix(choiceprice[,3:12]), 2, mean))

```

## Dispersion (variance) in prices of each product
```{r}
print(apply(as.matrix(choiceprice[,3:12]), 2, var))
```

## Average price of all products
```{r}
average = mean(unlist(choiceprice[,3:12]))
print(average)
```

## Dispersion (variance) in prices of all products
```{r}
dispersion = var(unlist(choiceprice[,3:12]))
print(dispersion)
```

## Market share of each product
```{r}
marketshare1 = data.frame(matrix(ncol = 10, nrow = 0))
colnames(marketshare1) = c(names(choiceprice[,3:12]))
for (i in 1:10){
  count = 0
  for (j in 1:nrow(choiceprice)){
    if(choiceprice[j,2]==i){
      count = count+1
    }
    marketshare1[1,i] = count/nrow(choiceprice)
  }
}
print(as.matrix(marketshare1))
```

## Market share by price bins
```{r}
marketshare2 = data.frame(matrix(ncol = 2, nrow = 0))
colnames(marketshare2) = c('Below average', 'Above average')
count = 0
for (i in 1:nrow(choiceprice)){
  if (choiceprice[i,choiceprice$choice[i]+2]<=average){
    count = count + 1
  }
}
marketshare2[1,1] = count/nrow(choiceprice)
marketshare2[1,2] = 1 - marketshare2[1,1]
print(as.matrix(marketshare2))
```

## Most popular choice for different types of households
```{r}
# Extract hhid for different types of households
incomeba = c()
incomeaa = c()
for (i in 1:nrow(demos)){
  if (demos$Income[i]<=mean(demos$Income)){
    incomeba = append(incomeba, demos$hhid[i])
  }
  else {
    incomeaa = append(incomeaa, demos$hhid[i])
  }
}
fs3orless = c()
fsmorethan3 = c()
for (i in 1:nrow(demos)){
  if (demos$Fam_Size[i]<=3){
    fs3orless = append(fs3orless, demos$hhid[i])
  }
  else {
    fsmorethan3 = append(fsmorethan3, demos$hhid[i])
  }
}
college = c()
nocollege = c()
for (i in 1:nrow(demos)){
  if (demos$college[i]==1){
    college = append(college, demos$hhid[i])
  }
  else {
    nocollege = append(nocollege, demos$hhid[i])
  }
}
whtcollar = c()
notwhtcollar = c()
for (i in 1:nrow(demos)){
  if (demos$whtcollar[i]==1){
    whtcollar = append(whtcollar, demos$hhid[i])
  }
  else {
    notwhtcollar = append(notwhtcollar, demos$hhid[i])
  }
}
retired = c()
notretired = c()
for (i in 1:nrow(demos)){
  if (demos$retired[i]==1){
    retired = append(retired, demos$hhid[i])
  }
  else {
    notretired = append(notretired, demos$hhid[i])
  }
}

# Define the function mode()
mode = function(x) {
  ux = unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Most popular choice by household types
popularchoice = data.frame(matrix(ncol = 10, nrow = 0))
colnames(popularchoice) = c('Income below average', 'Income above average',
                          'Family size 3 or less', 'Family size more than 3',
                          'College', 'No college', 'White collar',
                          'Not white collar', 'Retired', 'Not retired')
popularchoice[1,1] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,incomeba))$choice)
popularchoice[1,2] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,incomeaa))$choice)
popularchoice[1,3] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,fs3orless))$choice)
popularchoice[1,4] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,fsmorethan3))$choice)
popularchoice[1,5] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,college))$choice)
popularchoice[1,6] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,nocollege))$choice)
popularchoice[1,7] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,whtcollar))$choice)
popularchoice[1,8] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,notwhtcollar))$choice)
popularchoice[1,9] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,retired))$choice)
popularchoice[1,10] =
  mode(subset(choiceprice,is.element(choiceprice$hhid,notretired))$choice)
print(as.matrix(popularchoice))
```

# Exercise 2 First Model

```{r}
# Create the data "cp" by adding the variable "price" and "type" to the data
# "choiceprice", where "price" is the price of the product chosen and "type"
# is the type of the product chosen
cp = choiceprice
cp$price = NA
cp$type = NA
for (i in 1:nrow(cp)){
  cp$price[i] = cp[,cp$choice[i]+2][i]
  cp$type[i] = colnames(cp[cp$choice[i]+2])
}
cp[1:10,]
```

## Mixed logit model
* Choice probability $$p_{i}(j)=\frac{e^{c_{j}+\alpha*price_{ij}}}{\sum_{l}e^{c_{l}+\alpha*price_{il}}}$$
* Log likelihood function $$L(c,\alpha|Data=price)=\sum_{i=1}^{n}\sum_{l}log(p{i}(l))$$
* Maximum likelihood estimation $$\max_{c,\alpha}L(c,\alpha|Data)$$
```{r}
# Log likelihood function for the data "cp"
likelihood1 = function(par){
  cp$constant = 0
  cp$constant[cp$type=='PPk_Stk'] = par[1]
  cp$constant[cp$type=='PBB_Stk'] = par[2]
  cp$constant[cp$type=='PFl_Stk'] = par[3]
  cp$constant[cp$type=='PHse_Stk'] = par[4]
  cp$constant[cp$type=='PGen_Stk'] = par[5]
  cp$constant[cp$type=='PImp_Stk'] = par[6]
  cp$constant[cp$type=='PSS_Tub'] = par[7]
  cp$constant[cp$type=='PPk_Tub'] = par[8]
  cp$constant[cp$type=='PFl_Tub'] = par[9]
  pr = exp(cp$constant+par[10]*cp$price)/(exp(par[1]+par[10]*cp$PPk_Stk)
                                          +exp(par[2]+par[10]*cp$PBB_Stk)
                                          +exp(par[3]+par[10]*cp$PFl_Stk)
                                          +exp(par[4]+par[10]*cp$PHse_Stk)
                                          +exp(par[5]+par[10]*cp$PGen_Stk)
                                          +exp(par[6]+par[10]*cp$PImp_Stk)
                                          +exp(par[7]+par[10]*cp$PSS_Tub)
                                          +exp(par[8]+par[10]*cp$PPk_Tub)
                                          +exp(par[9]+par[10]*cp$PFl_Tub)
                                          +exp(par[10]*cp$PHse_Tub))
  return(-sum(log(pr)))
}

# Maximum likelihood estimation
est1 = optim(runif(10),likelihood1,method='BFGS')
```
The coefficient on price is `r est1$par[10]`. Therefore, an increase in the price of product j will decrease the probability that household i buys product j.

# Exercise 3 Second Model

```{r}
# Create the data "ci" by changing the variable "price" to "income"
ci = cp
colnames(ci)[13] = 'income'
ci$income = demos$Income[match(ci$hhid, demos$hhid)]
ci[1:10,]
```

## Mixed logit model
* Choice probability $$p_{i}(j)=\frac{e^{c_{j}+\beta*income_{i}}}{\sum_{l}e^{c_{l}+\beta*income_{i}}}$$
* Log likelihood function $$L(c,\alpha|Data=income)=\sum_{i=1}^{n}\sum_{l}log(p{i}(l))$$
* Maximum likelihood estimation $$\max_{c,\beta}L(c,\beta|Data)$$
```{r}
# Log likelihood function for the data "ci"
likelihood2 = function(par){
  ci$constant = 0
  ci$constant[ci$type=='PPk_Stk'] = par[1]
  ci$constant[ci$type=='PBB_Stk'] = par[2]
  ci$constant[ci$type=='PFl_Stk'] = par[3]
  ci$constant[ci$type=='PHse_Stk'] = par[4]
  ci$constant[ci$type=='PGen_Stk'] = par[5]
  ci$constant[ci$type=='PImp_Stk'] = par[6]
  ci$constant[ci$type=='PSS_Tub'] = par[7]
  ci$constant[ci$type=='PPk_Tub'] = par[8]
  ci$constant[ci$type=='PFl_Tub'] = par[9]
  pr = exp(ci$constant+par[10]*ci$income)/(exp(par[1]+par[10]*ci$income)
                                           +exp(par[2]+par[10]*ci$income)
                                           +exp(par[3]+par[10]*ci$income)
                                           +exp(par[4]+par[10]*ci$income)
                                           +exp(par[5]+par[10]*ci$income)
                                           +exp(par[6]+par[10]*ci$income)
                                           +exp(par[7]+par[10]*ci$income)
                                           +exp(par[8]+par[10]*ci$income)
                                           +exp(par[9]+par[10]*ci$income)
                                           +exp(par[10]*ci$income))
  return(-sum(log(pr)))
}

# Maximum likelihood estimation
est2 = optim(runif(10),likelihood2,method='BFGS')
```
The coefficient on income is `r est2$par[10]`. Therefore, an increase in household i's income will increase the probability that household i buys product j.

# Exercise 4 Marginal Effects

* The marginal effect of the price of product k on household i's demand of product j is $$\frac{\partial p{ij}}{\partial price_{ik}}=p_{ij}(\delta_{ijk}-p_{ik})\alpha=p_{ij}(\delta_{ijk}-p_{ik})(`r est1$par[10]`)$$, where $\delta_{ijk}$ is an indicator variable equal to 1 if $j=k$ and equal to 0 otherwise.

* The marginal effect of household i's income on household i's demand of product j is $$\frac{\partial p_{ij}}{\partial income_{i}}=p_{ij}(1-1)\beta=0$$.

# Exercise 5 IIA

## Full set of alternatives
```{r}
# Create the data "cf" by adding he variable "income" to the data "cp"
cf = cp
cf$income = ci$income
cf[1:10,]
```

### Mixed logit model
* Choice probability $$p_{i}(j)=\frac{e^{c_{j}+\alpha*price_{ij}+\beta*income_{i}}}{\sum_{l}e^{c_{l}+\alpha*price_{ij}+\beta*income_{i}}}$$
* Log likelihood function $$L(c,\alpha|Data=price,income)=\sum_{i=1}^{n}\sum_{l}log(p{i}(l))$$
* Maximum likelihood estimation $$\max_{c,\alpha,\beta}L(c,\alpha,\beta|Data)$$
```{r}
# Log likelihood function for the data "cf"
likelihood3 = function(par){
  cf$constant = 0
  cf$constant[cf$type=='PPk_Stk'] = par[1]
  cf$constant[cf$type=='PBB_Stk'] = par[2]
  cf$constant[cf$type=='PFl_Stk'] = par[3]
  cf$constant[cf$type=='PHse_Stk'] = par[4]
  cf$constant[cf$type=='PGen_Stk'] = par[5]
  cf$constant[cf$type=='PImp_Stk'] = par[6]
  cf$constant[cf$type=='PSS_Tub'] = par[7]
  cf$constant[cf$type=='PPk_Tub'] = par[8]
  cf$constant[cf$type=='PFl_Tub'] = par[9]
  pr = exp(cf$constant+par[10]*cf$price+par[11]*cf$income)/
    (exp(par[1]+par[10]*cf$PPk_Stk+par[11]*cf$income)
    +exp(par[2]+par[10]*cf$PBB_Stk+par[11]*cf$income)
    +exp(par[3]+par[10]*cf$PFl_Stk+par[11]*cf$income)
    +exp(par[4]+par[10]*cf$PHse_Stk+par[11]*cf$income)
    +exp(par[5]+par[10]*cf$PGen_Stk+par[11]*cf$income)
    +exp(par[6]+par[10]*cf$PImp_Stk+par[11]*cf$income)
    +exp(par[7]+par[10]*cf$PSS_Tub+par[11]*cf$income)
    +exp(par[8]+par[10]*cf$PPk_Tub+par[11]*cf$income)
    +exp(par[9]+par[10]*cf$PFl_Tub+par[11]*cf$income)
    +exp(par[10]*cf$PHse_Tub+par[11]*cf$income))
  return(-sum(log(pr)))
}

# Maximum likelihood estimation
est3 = optim(runif(11),likelihood3,method='BFGS')
```
The estimated coefficients on price and income are `r est3$par[10]` and `r est3$par[11]`, respectively.

## Subset of alternatives
```{r}
# Create the data “cr” by removing the choice “PImp_Stk” from the data “cf”
cr = cf[!(cf$type=='PImp_Stk'),]
cr[1:10,]
```

### Mixed logit model
```{r}
# Log likelihood function for the data "cr"
likelihood4 = function(par){
  cr$constant = 0
  cr$constant[cr$type=='PPk_Stk'] = par[1]
  cr$constant[cr$type=='PBB_Stk'] = par[2]
  cr$constant[cr$type=='PFl_Stk'] = par[3]
  cr$constant[cr$type=='PHse_Stk'] = par[4]
  cr$constant[cr$type=='PGen_Stk'] = par[5]
  cr$constant[cr$type=='PSS_Tub'] = par[6]
  cr$constant[cr$type=='PPk_Tub'] = par[7]
  cr$constant[cr$type=='PFl_Tub'] = par[8]
  pr = exp(cr$constant+par[9]*cr$price+par[10]*cr$income)/
    (exp(par[1]+par[9]*cr$PPk_Stk+par[10]*cr$income)
    +exp(par[2]+par[9]*cr$PBB_Stk+par[10]*cr$income)
    +exp(par[3]+par[9]*cr$PFl_Stk+par[10]*cr$income)
    +exp(par[4]+par[9]*cr$PHse_Stk+par[10]*cr$income)
    +exp(par[5]+par[9]*cr$PGen_Stk+par[10]*cr$income)
    +exp(par[6]+par[9]*cr$PSS_Tub+par[10]*cr$income)
    +exp(par[7]+par[9]*cr$PPk_Tub+par[10]*cr$income)
    +exp(par[8]+par[9]*cr$PFl_Tub+par[10]*cr$income)
    +exp(par[9]*cr$PHse_Tub+par[10]*cr$income))
  return(-sum(log(pr)))
}

# Maximum likelihood estimation
est4 = optim(runif(10),likelihood4,method='BFGS')
```
The estimated coefficients on price and income are `r est4$par[9]` and `r est4$par[10]`, respectively.

### Compute the test statistic:
```{r}
MTT = -2*(-likelihood4(est3$par[-6])+likelihood4(est4$par))
```
$MTT=`r MTT`$

### Conclude on IIA
```{r}
p = pchisq(MTT,df=10,lower.tail=FALSE)
```
$P(X>MTT|X\sim \chi^{2}(10))=`r p`$, so IIA is not violated.








