---
title: "Econ 613 Assignment 2"
author: "Xiaolin Ding"
date: "`r format(Sys.time(), '%m/%d/%y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
library(bayesm)
library(plyr)
library(dplyr)
library(fastDummies)
library(gmm)
library(tidyr)
library(tidyverse)
```

# Links to the datasets

```{r}
pop = read.csv("file:///Users/DXL/Desktop/Econ613/population.csv")
cri = read.csv("file:///Users/DXL/Desktop/Econ613/crime_long.csv")
off = read.csv("file:///Users/DXL/Desktop/Econ613/officers.csv")
```

# Exercise 2 Data Manipulation

* Total Crimes Per Month
```{r}
sum1 = cri %>% group_by(crime_month) %>%
  summarise(cripermon=sum(crimes))
month = as.Date(sum1 %>% pull(crime_month))
cripermon = sum1 %>% pull(cripermon)
plot(month,cripermon,col=1,type="l",xaxt='n',main='Total Crimes Per Month',
     xlab='crime_month',ylab='crimes',lwd=1)
axis.Date(1,at=seq(min(month),max(month),by='48 mon'),format='%Y-%m')
```

* Merged data df1
```{r}
df1 = full_join(cri, pop,
                    by=c('crime_month'='month','district'='district'))
df1[1:10,]
```

* Panel data df7 of unit over time with the following variables
  + Total crimes per resident
  + Violent crimes per resident
  + Property crimes per resident
  + Median income
  + Share of black, Hispanic, and white residents
```{r}
sum2 = df1 %>% group_by(crime_month,district) %>%
  summarise(cribydis=sum(crimes),popbydis=sum(tot_pop),white=sum(tot_white),
            black=sum(tot_black),hisp=sum(tot_hisp),
            med=quantile(p50_inc,1/2,na.rm=TRUE))
sum3 = df1 %>% group_by(crime_month,district,crime_type) %>%
  summarise(cribytype=sum(crimes))
popbydis = sum2 %>% pull(popbydis)
index1 = which(is.na(popbydis)==FALSE)
popbydis = popbydis[!is.na(popbydis)]
df2 = data.frame(matrix(ncol=0,nrow=length(popbydis)))
crimonth = sum2 %>% pull(crime_month)
df2$crimonth = crimonth[c(index1)]
dis = sum2 %>% pull(district)
df2$dis = dis[c(index1)]
cribydis = sum2 %>% pull(cribydis)
df2$cribydis = cribydis[c(index1)]
df2$popbydis = popbydis
white = sum2 %>% pull(white)
df2$white = white[c(index1)]
black = sum2 %>% pull(black)
df2$black = black[c(index1)]
hisp = sum2 %>% pull(hisp)
df2$hisp = hisp[c(index1)]
med = sum2 %>% pull(med)
df2$med = med[c(index1)]
crimon = sum3 %>% pull(crime_month)
df3 = data.frame(matrix(ncol=0,nrow=length(crimon)))
df3$crimon = crimon
df3$dis = sum3 %>% pull(district)
df3$type = sum3 %>% pull(crime_type)
df3$cribytype = sum3 %>% pull(cribytype)
df4 = full_join(df3, df2,
                    by=c('crimon'='crimonth','dis'='dis'))
index2 = which(is.na(df4$cribydis)==FALSE)
df5 = df4[c(index2),]
df5$criperres = df5$cribydis/df5$popbydis
df5$typeperres = df5$cribytype/df5$popbydis
df5$shareofb = df5$black/df5$popbydis
df5$shareofh = df5$hisp/df5$popbydis
df5$shareofw = df5$white/df5$popbydis
df6 = df5[,c(1:3,5,10:15)]
df6 = subset(df6, df6$type=='violent' | df6$type=='property')
df7 = spread(df6, type, typeperres)
colnames(df7) = c('month','district','total_crime','median_income',
                  'crime_per_resident','share_of_black',
                  'share_of_hispanic','share_of_white',
                  'property_per_resident','violent_per_resident')
df7[1:10,]
```

# Exercise 3 Panel Data: Introduction

* Create data df9 by merging the data 'off' with the previous panel data df7
```{r}
df8 = full_join(off, df7,
                    by=c('month'='month','unit'='district'))
df9 = df8[,c(1:7,9:11)]
df9 = df9[rowSums(is.na(df9[,1:10]))== 0,]
df9[1:10,]
```

* Estimate $\beta$ and $\gamma$
```{r}
X1 = as.matrix(df9[,c(4,6:10)])
Y1 = as.matrix(df9[,5])
coef1 = solve(t(X1)%*%X1)%*%t(X1)%*%Y1
coef1[1:6,1]
```
The estimated $\beta$ is `coef1[1,1]` and the estimated $\gamma$ are `r coef1[2:6,1]`.

# Exercise 4 Panel Data: More controls

* Create data df10 with district and month dummies
```{r}
df10 = dummy_cols(df9, select_columns=c('unit','month'))
df10[1:10,]
```

* Estimate $\beta$, $\gamma$, $\psi$, and $\kappa$ with district and year and month dummies, whereas the base group dummies are excluded
```{r}
X2 = as.matrix(df10[,c(4,6:34,36:166)])
Y2 = as.matrix(df10[,5])
coef2 = solve(t(X2)%*%X2)%*%t(X2)%*%Y2
coef2[1:6,1]
```
The estimated $\beta$ is `r coef2[1,1]`, the estimated $\gamma$ are `r coef2[2:6,1]`, the estimated $\psi$ are `r coef2[7:30,1]`, and the estimated $\kappa$ are `r coef2[31:161,1]`.

# Exercise 5 Panel Data: Individual fixed effects

* Create data df11 with mean differences, df12 with means, and df13 with first differences
```{r}
df11 = df10 %>% group_by(NUID) %>%
  mutate(across(!month & !unit, ~.x-mean(.x)))
df12 = df10 %>% group_by(NUID) %>%
  summarise(across(!month & !unit, list(mean)))
df13 = df10 %>% group_by(NUID) %>%
  mutate(across(!month & !unit, ~.x-lag(.x)))
df13 = df13[rowSums(is.na(df13[,1:10]))== 0,]
df11 = as.data.frame(df11)
df12 = as.data.frame(df12)
df13 = as.data.frame(df13)
df11[1:10,]
df12[1:10,]
df13[1:10,]
```

* Within estimator
```{r}
X3 = as.matrix(df11[,c(4,6:34,36:166)])
Y3 = as.matrix(df11[,5])
coef3 = solve(t(X3)%*%X3)%*%t(X3)%*%Y3
coef3[1,1]
```
The estimated $\beta$ is `r coef3[1,1]`.

* Between estimator
```{r}
X4 = as.matrix(df12[,c(2,4:32,34:164)])
Y4 = as.matrix(df12[,3])
coef4 = solve(t(X4)%*%X4,tol=1e-100)%*%t(X4)%*%Y4
coef4[1,1]
```
The estimated $\beta$ is `r coef4[1,1]`. Here taking the mean does not eliminate individual fixed effects, but since there are too many individuals, the calculation amount is too large if we estimate the individual fixed effects for each individual.

* First difference estimator
```{r}
X5 = as.matrix(df13[,c(4,6:34,36:166)])
Y5 = as.matrix(df13[,5])
coef5 = solve(t(X5)%*%X5,tol=1e-100)%*%t(X5)%*%Y5
coef5[1,1]
```
The estimated $\beta$ is `r coef5[1,1]`. The within and between estimators are negative with small absolute value, while the first difference estimator is positive with small absolute value.

* GMM
  + Estimate
  $$
  min_{\beta}[(Y-X\beta)^{'}X]I[X^{'}(Y-X\beta)]
  $$
  + Obtain
  $$
  \hat{\beta_{1}}=(X^{'}XX^{'}X)^{-1}X^{'}XX^{'}Y
  $$
```{r}
coef6 = solve(t(X2)%*%X2%*%t(X2)%*%X2,tol=1e-100)%*%t(X2)%*%X2%*%t(X2)%*%Y2
coef6[1:6,1]
```

  + Matrix of variance/covariance
  $$\hat{\mathbb{S}}_{n}=\frac{1}{N}X^{'}(Y-X\hat{\beta_{1}})(Y-X\hat{\beta_{1}})^{'}X
  $$
```{r}
s = t(X2)%*%(Y2-X2%*%coef6)%*%t(Y2-X2%*%coef6)%*%X2/nrow(X2)
s[1:6,1:6]
```

  + Solve the problem
  $$
  min_{\beta}[(Y-X\beta)^{'}X]\mathbb{S_{n}}[X^{'}(Y-X\beta)]
  $$
  + Obtain
  $$
  \hat{\beta}_{GMM}=(X^{'}X\mathbb{S_{n}}X^{'}X)^{-1}X^{'}X\mathbb{S_{n}}X^{'}Y
  $$
```{r}
coef7 = solve(t(X2)%*%X2%*%s%*%t(X2)%*%X2,tol=1e-100)%*%
  t(X2)%*%X2%*%s%*%t(X2)%*%Y2
coef7[1:6,1]
```
The estimated $\beta$ is `r coef7[1,1]`, the estimated $\gamma$ are `r coef7[2:6,1]`, the estimated $\psi$ are `r coef7[7:30,1]`, and the estimated $\kappa$ are `r coef7[31:161,1]`.





